---
title: "Survival Analysis HW1"
author: "313657003周佳萱"
date: "`r Sys.Date()`"
output:
  pdf_document:
    latex_engine: xelatex
header-includes:
- \usepackage{fontspec}
- \usepackage{xeCJK}
- \usepackage{amsmath}
- \usepackage{amssymb}
- \setCJKmainfont{新細明體}
mathjax: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1.
## (a)

\begin{align*}
S(t) &= \exp\left(-\int^{t}_{0} \lambda (u) \, du\right) = \exp\left(-\int^{t}_{0} c \, du\right) = e^{-ct} \\
f(t) &= -\frac{d S(t)}{dt}= -\frac{d e^{-ct}}{dt} = c e^{-ct}
\end{align*}



### Median failure time for c=2, 5, 11

Let $t_{\eta}$ is median failure time,


\begin{align*}
0.5 &= 1 - e^{-ct_{\eta}} \\
-\log 2 &= -ct_{\eta} \\
t_{\eta} &= \frac{\log 2}{c}
\end{align*}



```{r median}
c = c(2,5,11) 
t = log(2) / c
t
```


$c = 2$, $t_{\eta} = \frac{\log 2}{2} = 0.3466$  
$c = 5$, $t_{\eta} = \frac{\log 2}{5} = 0.1386$  
$c = 11$, $t_{\eta} = \frac{\log 2}{11} = 0.0630$ 


## (b)

\begin{align*}
f(t) &= -\frac{d S(t)}{dt} = -\frac{d \exp(- \theta t^{\beta} )}{dt} 
     = \theta \beta t^{\beta -1} \exp(- \theta t^{\beta}) \\
\lambda (t) &= \frac{f(t)}{S(t)} = \frac{\theta \beta t^{\beta -1} \exp(- \theta t^{\beta})}{\exp(- \theta t^{\beta} ) } 
           =  \theta \beta t^{\beta -1} 
\end{align*}


# Problem 2.
## (a) Calculate the empirical survival function and graph it.

```{r fig.width=10, fig.height=7, fig.align="center", warning=FALSE}

library(knitr)

survival_times <- c(7, 47, 58, 74, 177, 232, 273, 285, 317, 429, 440,
                    445, 455, 468, 495, 497, 532, 571, 579, 581, 650, 702,
                    715, 779, 881, 900, 903, 968, 1077, 1109, 1314, 1334,
                    1367, 1534, 1712, 1784, 1877, 1886, 2045, 2056, 2260,
                    2429, 2509)

n <- length(survival_times)

survival_prob <- ((n-1):0) / n  

survival_table_individual <- data.frame(Survival_Time = survival_times, Survival_Prob = survival_prob)

kable(survival_table_individual, align = "c")

plot(survival_times, survival_prob, pch = 16, col = "black")   
points(survival_times[-1], ((n-1):1) / n, pch = 1, col = "black")  
segments(survival_times[-n], survival_prob[-n], survival_times[-1], ((n-1):1) / n
         , col = "black", lwd = 2)
grid()
```

## (b) Estimate the 1-year, 2-year, 3-year, 4-year and 5-year survival rates by nonparametric methods.

```{r Survival Rate, warning=FALSE}

survival_table <- data.frame(Year = 1:5, Rate = rep(0, 5))

# Survival Rate
for(i in 1:5){
  survival_table$Rate[i] <- length(survival_times[survival_times > 365 * i]) / n
}


library(knitr)

kable(survival_table
      , caption = "Estimated Survival Rates"
      , align = "c")

```

##  (c) Under nonparametric model, construct 95% con dence intervals for the 1-year, 2-year, 3-year, 4-year and 5-year survival rates.

```{r Confidence Interval}

survival_table$LowerCI <- rep(0, 5)
survival_table$UpperCI <- rep(0, 5)

for(i in 1:5){
  sur_exp <- survival_table$Rate[i]  
  sur_var <- (sur_exp * (1 - sur_exp)) / n 
  
  survival_table$LowerCI[i] <- sur_exp - (1.96 * sqrt(sur_var))
  survival_table$UpperCI[i] <- sur_exp + (1.96 * sqrt(sur_var))
}

survival_table <- survival_table[, c("Year", "LowerCI", "Rate", "UpperCI")]

kable(survival_table, caption = "Estimated Survival Rates with 95% Confidence Interval", align = "c")
```


## (d) Estimate the median survival time.

According to Table 1 in (a), we can see that $S(650) = 0.5116279$ and $S(702) = 0.4883721$, so the median survival time is 650.


## (e) Under the exponential model with the constant hazard $\lambda$, calculate the maximum likelihood estimate of $\lambda$ and construct a $95\%$ confidence interval for $\lambda$.

\begin{align*}
X &\sim \text{Exp}(\lambda), \quad f(x \mid \lambda) = \lambda e^{-\lambda x} \\
&\Rightarrow \mathcal{L} (\lambda \mid x) = \prod_{i=1}^{43} f(x_i \mid \lambda) = \lambda^{43} e^{-\lambda \sum_{i=1}^{43} x_i} \\
&\Rightarrow  \log \mathcal{L} (\lambda \mid x) = 43 \log \lambda - \lambda \sum_{i=1}^{43} x_i \\
&\Rightarrow \frac{\partial}{\partial \lambda} \log \mathcal{L} (\lambda \mid x) = \frac{43}{\lambda} - \sum_{i=1}^{43} x_i \equiv 0 \\
&\Rightarrow \hat{\lambda} = \frac{43}{\sum_{i=1}^{43} x_i} \\
&\Rightarrow \text{The maximum likelihood of} \, \lambda \,  \text{is} \,  \frac{43}{\sum_{i=1}^{43} x_i} \\
   \sum_{i=1}^{43} X_i &\sim \Gamma(43, \lambda) \\
&\Rightarrow 2\lambda \sum_{i=1}^{43} X_i \sim \Gamma(43, \frac{1}{2}) \equiv \chi^2(86) \\
&\Rightarrow 0.95 = P(\chi^2_{0.025}(86) \leq 2\lambda \sum_{i=1}^{43} x_i \leq \chi^2_{0.975}(86)) \\
&\Rightarrow 0.95 = P\left( \frac{\chi^2_{0.025}(86)}{2 \sum_{i=1}^{43} x_i} \leq \lambda \leq \frac{\chi^2_{0.975}(86)}{2 \sum_{i=1}^{43} x_i} \right) 
\end{align*}

```{r confidence interval}

chi_low <- qchisq(0.025, df = 86)
chi_high <- qchisq(0.975, df = 86)

sum_survival <- sum(survival_times)

LowerCI <- chi_low / (2 * sum_survival)
UpperCI <- chi_high / (2 * sum_survival)

cat("95% Confidence Interval for Lambda:\n")
cat("Lower Bound:", LowerCI, "\n")
cat("Upper Bound:", UpperCI, "\n")

```

$95\%$ Confidence Interval for $\lambda$ is $(0.000783, 0.001428)$

## (f)Do you think the exponential model is appropriate for the data? Explain.

```{r exponential model, fig.width=10, fig.height=7, fig.align="center", warning=FALSE}

# MLE
mle <- 43 / sum_survival  

exp_fun <- function(x) {
  exp(-mle * x)
}

x <- 0:2500  
exp_val <- exp_fun(x)

# Exponential 
plot(x, exp_val, type = "l", col = "red", lwd = 2, 
     xlab = "Time", ylab = "Survival Probability",
     main = "Comparison of Exponential and Empirical Survival Function")

# Original Survival
points(survival_times, survival_prob, pch = 16, col = "black")  
points(survival_times[-1], ((n-1):1) / n, pch = 1, col = "black") 
segments(survival_times[-n], survival_prob[-n], survival_times[-1], ((n-1):1) / n
         , col = "black", lwd = 2)

grid()

legend("topright", legend = c("Exponential Model", "Empirical Data"),
       col = c("red", "black"), lty = c(1, NA), pch = c(NA, 16), bty = "n")

```


Compared the exponential survival function with the empirical survival function, we observe that they have a similar shape. Therefore, the exponential model is  appropriate for this data.

# Problem 3


$$\begin{aligned}
m(t)&=E(T-t\mid T>t)\\
&=\int_t^\infty (T-t)\frac{f(T)}{S(t)}dT\\
&=\int_t^\infty T\frac{f(T)}{S(t)}dT - \int_t^\infty \frac{tf(T)}{S(t)}dT\\
&=\frac{1}{S(t)}   [     \int_t^\infty Tf(T)dT  ] -t
\end{aligned}$$


$$\begin{aligned}
m^\prime(t)
&= \frac{-tf(t)S(t)+f(t) \int_t^\infty T f(T)dT}{S(t)^2}\\
&= \frac{f(t)}{S^2(t)}\int_t^\infty Tf(T)dT-\frac{tf(t)}{S(t)}-1
\end{aligned}$$

then, we can make the relationship between $m(t)$ and $m^\prime(t)$

$$\begin{aligned}
\frac{m(t)f(t)}{S(t)} 
&= \frac{f(t) \int_t^\infty T f(T)dT}{S(t)^2} - \frac{tf(t)}{S(t)}\\
\Rightarrow & m^\prime(t) = \frac{m(t)f(t)}{S(t)} -1 \\
\Rightarrow &m^\prime(t)+1 = \frac{f(t)}{S(t)}m(t)\\
\Rightarrow & -\frac{S^\prime(t)}{S(t)}=\frac{m^\prime(t)}{m(t)}+\frac{1}{m(t)}\\
\Rightarrow & -\log S(t)=\log m(t)+\int_0^t \frac{1}{m(u)}du+c\\
\Rightarrow & S(t)=\frac{1}{m(t)}\exp\left(-\int_0^t\frac{1}{m(u)}du-c\right)\\
\Rightarrow & 1=S(0)=\frac{1}{m(0)}e^{-c}\Rightarrow\ c=-\log m(0)\\
\end{aligned}$$

$$ \therefore S(t)
=\frac{1}{m(t)}\exp\left(-\int_0^t\frac{1}{m(u)}du + \log m(0) \right) \\
=\frac{m(0)}{m(t)}\exp\left(-\int_0^t\frac{1}{m(u)}du\right)$$


# Problem 4
##  (a) Calculate both the Kaplan-Meier estimate and the empirical survival distribution on the basis of the placebo data. What is the relationship between the Kaplan-Meier estimate and the empirical survival distribution from your calculation?

```{r placebo KM and empirical}
library(knitr)

placebo <- c(1, 1, 2, 2, 3, 4, 4, 5, 5, 8, 8, 8, 8, 11, 11, 12, 12, 15, 17, 22, 23)
n_placebo <- length(placebo)

time_points <- sort(unique(placebo))

placebo_survival_table <- data.frame(
  Time = time_points,
  KM_Estimate = numeric(length(time_points)),
  Empirical_Survival = numeric(length(time_points))
)

last_survival <- 1

#  KM and Empirical Survival
for (i in seq_along(time_points)) {
  t <- time_points[i]
  d <- sum(placebo == t)
  n_risk <- sum(placebo >= t)
  
  s_t <- last_survival * (1 - d / n_risk)
  placebo_survival_table[i, "KM_Estimate"] <- s_t
  last_survival <- s_t
  
  placebo_survival_table[i, "Empirical_Survival"] <- sum(placebo > t) / n_placebo
}


kable(
  placebo_survival_table,
  caption = "Kaplan-Meier Estimate and Empirical Survival Distribution for Placebo Group",
  align = "c",
)

```

There is no difference between KM estimator and empirical survival function on placebo group.

## (b) Do you observe a similar phenomenon which also holds for other uncensored survival data? Explain.

Under uncensored data, empirical survival function is same as Kaplan-Meier estimator.

$$\begin{aligned}
\hat{S}_{KM}(t)&=\prod_{y_{(j)\le t}}\left(1-\frac{d_{(j)}}{N_{(j)}}\right)\\
&=\prod_{y_{(j)\le t}}\frac{\sum_{i=1}^{N} I(y_i>y_{(j)})}{\sum_{i=1}^{N}I(y_i\ge y_{(j)})}\\
&=\prod_{y_{(j)\le t}}\frac{\sum_{i=1}^{N} I(y_i>y_{(j)})/N}{\sum_{i=1}^{N} I(y_i>y_{(j-1)})/N}\quad
\because \text{uncensored}\\
&=\prod_{y_{(j)\le t}}\frac{\hat{S}_e (y_{(j)})}{\hat{S}_e (y_{(j-1)})}\\
&=\frac{\hat{S}_e (y_{(1)})}{\hat{S}_e (y_{(0)})}\times \cdots\times\frac{\hat{S}_e (y_{(i-1)})}{\hat{S}_e (y_{(i-2)})}\\
&=\frac{\hat{S}_e (y_{(i-1)})}{\hat{S}_e (y_{(0)})}\\
&=\hat{S}_e (y_{(i-1)})\\
&=\sum_{i=1}^N \frac{I(y_s>y_{(i-1)})}{N}\\
&=\hat{S}_e (t)
\end{aligned}$$


## (c) Calculate the Kaplan-Meier estimate on the basis of the 6MP data.


```{r placebo KM and empirical 6MP}

MP<-c(6, 6, 6, 7, 10, 13, 16, 22, 23,6, 9, 10, 11, 17, 19, 20, 25, 32, 32, 34, 35)
unsensor_MP <- c(6, 6, 6, 7, 10, 13, 16, 22, 23)

n_MP <- length(MP)

time_points <- sort(unique(unsensor_MP))

MP_survival_table <- data.frame(
  Time = time_points,
  KM_Estimate = numeric(length(time_points))
)

last_survival <- 1

# KM and Empirical Survival
for (i in seq_along(time_points)) {
  t <- time_points[i]
  d <- sum(unsensor_MP == t)
  n_risk <- sum(MP >= t)
  
  s_t <- last_survival * (1 - d / n_risk)
  MP_survival_table[i, "KM_Estimate"] <- s_t
  last_survival <- s_t
}


kable(
  MP_survival_table,
  caption = "Kaplan-Meier Estimate for MP Group",
  align = "c",
)

```



## (d) Use the variance estimates that you learn from class to produce a 95% condence interval at each observed failure time for the empirical survival distribution on the basis of the placebo data. Comment on your results.


```{r Confidence Ingterval Placebo empirical}

placebo_survival_table$LowerCI <- rep(0, nrow(placebo_survival_table))
placebo_survival_table$UpperCI <- rep(0, nrow(placebo_survival_table))

for(i in 1:nrow(placebo_survival_table)){

  sur_exp <- placebo_survival_table$Empirical_Survival[i]  
  sur_var <- (sur_exp * (1 - sur_exp)) / n_placebo
  sur_low <- sur_exp - (1.96 * sqrt(sur_var))
  sur_upper <- sur_exp + (1.96 * sqrt(sur_var))
  
  placebo_survival_table$LowerCI[i] <- ifelse(sur_low <= 0, 0, sur_low)
  placebo_survival_table$UpperCI[i] <- ifelse(sur_upper >= 1, 1, sur_upper)
}

placebo_survival_table_CI <- placebo_survival_table[, c("Time", "LowerCI", "Empirical_Survival", "UpperCI")]

kable(placebo_survival_table_CI 
      , caption = "Empirical Survival with 95% Confidence Interval (Placebo Group)"
      , align = "c")
```

When calculating the 95% confidence interval for each empirical survival point, we observe that some lower bounds fall below 0. It occurs because the sample size is relatively small (only 21 data points), making the normal approximation less accurate in the tails.


## (e) Use the variance estimates that you learn from class to produce a 95% condence interval, at each uncensored time, for the the Kaplan-Meier estimate on the basis of the 6MP data. Comment on your results.


```{r placebo KM and empirical 6MP CI}

time_points <- sort(unique(unsensor_MP))

MP_survival_table$LowerCI <- rep(0,length(time_points))
MP_survival_table$UpperCI <- rep(0,length(time_points))

temp <- 0

for (i in seq_along(time_points)) {
  t <- time_points[i]
  
  d <- sum(unsensor_MP == t)
  
  n_risk <- sum(MP >= t)
  
  temp <- temp + d/(n_risk*(n_risk-d))
  km <- MP_survival_table$KM_Estimate[i]
  
  sur_var <- km^2 * temp
  
  lower<-km-1.96*sqrt(sur_var)
  upper<-km+1.96*sqrt(sur_var)

  MP_survival_table$LowerCI[i] <- ifelse(lower < 0, 0, lower)
  MP_survival_table$UpperCI[i] <- ifelse(upper > 1, 1, upper)
  
}

MP_survival_table_CI <- MP_survival_table[, c("Time", "LowerCI", "KM_Estimate", "UpperCI")]


kable(
  MP_survival_table_CI,
  caption = "Kaplan-Meier Estimate with 95% Confidence Interval (6MP Group)",
  align = "c",
)

```

Compared to part (d), the confidence intervals are less likely to go beyond [0,1]. Still, with only 21 data points, the intervals widen over time.

## (f) Compare the survival function estimates from the placebo and 6MP groups. Interpret your results.

```{r plot KM for placebo and 6MP}

plot(0, 0, type = "n", xlim = c(0, max(MP_survival_table$Time, placebo_survival_table$Time)),
     ylim = c(0, 1), xlab = "Time", ylab = "Survival Probability",
     main = "Kaplan-Meier Survival Curves: 6MP vs Placebo")

# --- 6MP Group ---

points(c(0, MP_survival_table$Time[-nrow(MP_survival_table)])
       , c(1, MP_survival_table$KM_Estimate[-nrow(MP_survival_table)])
       , pch = 16, col = "red")

points(MP_survival_table$Time,
       c(1, MP_survival_table$KM_Estimate[-nrow(MP_survival_table)]),
       pch = 1, col = "red")

segments(x0 = c(0, MP_survival_table$Time[-nrow(MP_survival_table)]),
         y0 = c(1, MP_survival_table$KM_Estimate[-nrow(MP_survival_table)]),
         x1 = MP_survival_table$Time,
         y1 = c(1, MP_survival_table$KM_Estimate[-nrow(MP_survival_table)]),
         col = "red")

# --- Placebo Group ---

points(c(0, placebo_survival_table$Time[-nrow(placebo_survival_table)])
       , c(1, placebo_survival_table$KM_Estimate[-nrow(placebo_survival_table)])
       , pch = 16, col = "blue")
points(placebo_survival_table$Time,
       c(1, placebo_survival_table$KM_Estimate[-nrow(placebo_survival_table)]),
       pch = 1, col = "blue")

segments(x0 = c(0, placebo_survival_table$Time[-nrow(placebo_survival_table)]),
         y0 = c(1, placebo_survival_table$KM_Estimate[-nrow(placebo_survival_table)]),
         x1 = placebo_survival_table$Time,
         y1 = c(1, placebo_survival_table$KM_Estimate[-nrow(placebo_survival_table)]),
         col = "blue")
grid()

legend("topright", legend = c("6MP", "Placebo"),
       col = c("red", "blue"), pch = 16, lty = 1, bty = "n")

```

From the graph above, we can see that the survival function of the 6MP treatment group is  higher than that of the placebo group.
It indicates that leukemia patients who received 6MP treatment had a higher survival rate compared to those who were given a placebo.

# Problem 5. Simulate survival data and compute the Nelson-Aalen estimator.
## (a) Simulate 100 event times with constant hazard 0.9 (using rexp()). Also simulate 100 censoring times with uniform distribution on [0,5]. Next, compute the observed times and an event indicator. How many observed events/censorings do you have?


```{r NA1}

set.seed(123)

event_time <- rexp(100, 0.9) #T
censor_time <- runif(100, 0, 5) #C

observed_time <- pmin(event_time, censor_time)   #Y  
event_indicator <- as.numeric(event_time <= censor_time) #I


n_event <- sum(event_indicator)       
n_censor <- sum(event_indicator == 0) 


cat("Number of observed events:", n_event, "\n")
cat("Number of censored observations:", n_censor, "\n")

```


## (b) compute Nelson-Aalen (cumulative hazard).

```{r NA2}

data <- data.frame(time = observed_time, status = event_indicator)

event_times <- sort(unique(data$time[data$status == 1]))

hazard_table <- data.frame(
  time = event_times,
  d = numeric(length(event_times)),
  n_risk = numeric(length(event_times)),
  d_over_n = numeric(length(event_times)),
  cum_hazard = numeric(length(event_times)),
  survival = numeric(length(event_times))
)


cum_hazard <- 0
for (i in seq_along(event_times)) {
  t <- event_times[i]
  d_j <- sum(data$time == t & data$status == 1)
  n_j <- sum(data$time >= t)
  d_over_n <- d_j / n_j
  cum_hazard <- cum_hazard + d_over_n
  
  hazard_table$d[i] <- d_j
  hazard_table$n_risk[i] <- n_j
  hazard_table$d_over_n[i] <- d_over_n
  hazard_table$cum_hazard[i] <- cum_hazard
  hazard_table$survival[i] <- exp(-cum_hazard)
}

kable(
  head(hazard_table[, c("time", "cum_hazard", "survival")], 10),
  caption = "Nelson-Aalen Cumulative Hazard Estimate (First 10 Events)",
  align = "c",
)

```



## (c) Plot the Nelson-Aalen estimator. Compare it to the true cumulative hazard using curve. For comparison, you may want to restrict plotting to time regions where there is more data. E.g., restrict to the true median of the survival distribution.


$$\begin{aligned}
f(t)  &= 0.9exp(-0.9t) \\
S(t)  &= exp(-0.9t)   \\
\therefore \Lambda(t)&=0.9t
\end{aligned}$$

True cumulative hazard is $\Lambda(t)=0.9t$


```{r NAplot}

max_time <- max(hazard_table$time)

plot(0, 0, xlab = "Time", ylab = "Cumulative Hazard",
     main = "Nelson-Aalen vs True Cumulative Hazard",
     xlim = c(0, max_time), ylim = c(0, 0.9 * max_time))

points(hazard_table$time, hazard_table$cum_hazard, col = "blue", lwd = 2,)
curve(0.9 * x, from = 0, to = max_time, col = "red", lty = 2, lwd = 2, add = TRUE)


legend("topleft", legend = c("Nelson-Aalen Estimate", "True Cumulative Hazard"),
       col = c("blue", "red"), lty = c(1, 2), lwd = 2, bty = "n")


```
When t < 1.5, Nelson-Aalen estimate closely follows the true cumulative hazard line.
When t > 1.5, Nelson-Aalen estimate begins to deviate from the true curve.



# Problem 6.
## (a) Reproduce simulation study in Updated Survival Notes, Second Chapter, Page30. (Attached Code)

```{r reproduce page 30}

set.seed(123)

n_list <- c(100, 10000)
lmbd <- 0.1
beta_list <- c(10, 5)
label_matrix <- matrix(c("KM1", "KM2", "KM3", "KM4"), nrow = 2)

true_surv <- function(t) exp(-lmbd * t)

KM1 <- KM2 <- KM3 <- KM4 <- NULL

for (ni in seq_along(n_list)) {
  n <- n_list[ni]
  event_time <- rexp(n, rate = lmbd)
  
  for (bi in seq_along(beta_list)) {
    beta <- beta_list[bi]
    label <- label_matrix[ni, bi]
    
    censor_time <- runif(n, 0, beta)
    observed_time <- pmin(event_time, censor_time)
    event_indicator <- as.numeric(event_time <= censor_time)
    data <- data.frame(time = observed_time, status = event_indicator)
    
    time_points <- sort(unique(data$time[data$status == 1]))
    
    KM_df <- data.frame(
      time = time_points,
      true = true_surv(time_points),
      KM = numeric(length(time_points))
    )
    
    km <- 1
    for (j in seq_along(time_points)) {
      t <- time_points[j]
      d <- sum(data$time == t & data$status == 1)
      n_risk <- sum(data$time >= t)
      km <- km * (1 - d / n_risk)
      KM_df$KM[j] <- km
    }
    
    assign(label, KM_df)
  }
}

par(mfrow = c(1, 2))  # 1 列 2 欄

# ----- Left：n = 100 -----
plot(0, 0, type = "n",
     xlim = c(0, 12), ylim = c(0, 1),
     xlab = "Time", ylab = "Survival Function",
     main = "n = 100")


curve(true_surv(x), from = 0, to = 12, col = "black", lwd = 3, add = TRUE)

lines(KM1$time, KM1$KM, col = "red", lwd = 2)
lines(KM2$time, KM2$KM, col = "blue", lwd = 2)

legend("topright",
       legend = c("True", "KM1 (β=10)", "KM2 (β=5)"),
       col = c("black", "red", "blue"),
       lty = 1, lwd = c(3,2,2), bty = "n")

# ----- Right：n = 10000 -----
plot(0, 0, type = "n",
     xlim = c(0, 12), ylim = c(0, 1),
     xlab = "Time", ylab = "Survival Function",
     main = "n = 10000")

curve(true_surv(x), from = 0, to = 12, col = "black", lwd = 3, add = TRUE)

lines(KM3$time, KM3$KM, col = "red", lwd = 2)
lines(KM4$time, KM4$KM, col = "blue", lwd = 2)

legend("topright",
       legend = c("True", "KM3 (β=10)", "KM4 (β=5)"),
       col = c("black", "red", "blue"),
       lty = 1, lwd = c(3,2,2), bty = "n")

```


## (b) Regenerate 100 simulation studies in (a). Based on these 100 simulation results, plot the average of estimated survival curve and 95% condence interval. (Attached Code)

```{r reproduce page 30 b}

set.seed(123)

n <- 100
lmbd <- 0.1
beta_list <- c(10, 5)
label_list <- c("KM1", "KM2")

for (bi in seq_along(beta_list)) {
  beta <- beta_list[bi]
  label <- label_list[bi]

  event_time <- rexp(n, rate = lmbd)
  censor_time <- runif(n, 0, beta)
  observed_time <- pmin(event_time, censor_time)
  status <- as.numeric(event_time <= censor_time)

  time_points <- sort(unique(observed_time[status == 1]))
  KM_df <- data.frame(
    time = time_points,
    KM = numeric(length(time_points)),
    LowerCI = numeric(length(time_points)),
    UpperCI = numeric(length(time_points))
  )

  km <- 1
  temp <- 0
  for (j in seq_along(time_points)) {
    t <- time_points[j]
    d <- sum(observed_time == t & status == 1)
    n_risk <- sum(observed_time >= t)

    km <- km * (1 - d / n_risk)
    KM_df$KM[j] <- km

    
    temp <- temp + d / (n_risk * (n_risk - d))
    var <- km^2 * temp
    KM_df$LowerCI[j] <- max(0, km - 1.96 * sqrt(var))
    KM_df$UpperCI[j] <- min(1, km + 1.96 * sqrt(var))
  }

  assign(label, KM_df)
}

par(mfrow = c(1, 2))  

# ----- Left：beta = 10 -----
plot(0, 0, type = "n",
     xlim = c(0, max(KM1$time)), ylim = c(0, 1),
     xlab = "Time", ylab = "Survival Function",
     main = "beta = 10")

polygon(c(KM1$time, rev(KM1$time)),
        c(KM1$UpperCI, rev(KM1$LowerCI)),
        col = rgb(0.6, 0.8, 1, 0.3), border = NA)

lines(KM1$time, KM1$KM, col = "red", lwd = 3)
lines(KM1$time, KM1$LowerCI, col = "blue", lwd = 2.5, lty = 2)
lines(KM1$time, KM1$UpperCI, col = "blue", lwd = 2.5, lty = 2)


curve(exp(-0.1 * x), from = 0, to = max(KM1$time), add = TRUE,
      col = "black", lwd = 2.5, lty = 3)

legend("bottomleft",
       legend = c("KM Curve", "95% CI", "True Survival"),
       col = c("red", "blue", "black"),
       lty = c(1, 2, 3),
       lwd = c(3, 2.5, 2.5),
       bty = "n")


# ----- Right：beta = 5 -----
plot(0, 0, type = "n",
     xlim = c(0, max(KM2$time)), ylim = c(0, 1),
     xlab = "Time", ylab = "Survival Function",
     main = "beta = 5")

polygon(c(KM2$time, rev(KM2$time)),
        c(KM2$UpperCI, rev(KM2$LowerCI)),
        col = rgb(0.6, 0.8, 1, 0.3), border = NA)

lines(KM2$time, KM2$KM, col = "red", lwd = 3)
lines(KM2$time, KM2$LowerCI, col = "blue", lwd = 2.5, lty = 2)
lines(KM2$time, KM2$UpperCI, col = "blue", lwd = 2.5, lty = 2)

curve(exp(-0.1 * x), from = 0, to = max(KM2$time), add = TRUE,
      col = "black", lwd = 2.5, lty = 3)

legend("bottomleft",
       legend = c("KM Curve", "95% CI", "True Survival"),
       col = c("red", "blue", "black"),
       lty = c(1, 2, 3),
       lwd = c(3, 2.5, 2.5),
       bty = "n")


```







